### 2023-02-07 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_interpreting_probabilities@2c56085dbe296d98270202ca1acb3669231eb772 🚀

\hypertarget{overview-of-interaction-probabilities}{%
\subsection{Overview of interaction
probabilities}\label{overview-of-interaction-probabilities}}

How are interaction probabilities defined in the literature? It might
not be as intuitive as one would think.

\begin{itemize}
\tightlist
\item
  There is a big difference in how we interpret the probability that two
  species \emph{can} interact (metaweb) and the probability that they
  \emph{will} interact (realized networks).
\item
  Interaction probabilities can be used to describe Boolean interactions
  (e.g., the probability that two species interact) and weighted
  interactions (e.g., the probability distribution of the amount of
  energy flow between two species).
\item
  In many studies, it is not obvious if authors use interaction scores
  or probabilities (in the latter case, it is rarely specified what
  these probabilities truly represent).
\end{itemize}

Papers: Guimarães (2020)

@@

\begin{itemize}
\tightlist
\item
  It means the probability that two taxa can interact, regardless of
  biological plasticity, environmental variability, or co-occurrence.
\item
  One observation is enough to set this probability to one.
\item
  Can we turn this into a local network realisation that is also
  probabilistic and intuitive?
\end{itemize}

@@

@@

@@

>>>>

The basic unit of food webs and other ecological networks are
individuals that interact with each others (e.g., by predation; Elton
2001). The aggregation of these individuals into more or less
homogeneous groups (e.g., populations, species, trophic species) allows
us to represent networks at different scales, which impacts the
properties and behaviour of these systems (Guimarães 2020). A network's
nodes can thus designate distinct levels of organization, whereas the
edges linking these nodes can describe a variety of interaction
measures. When using a Boolean (yes-no) representation of biotic
interactions, the observation that one individual from group (or node)
\(i\) interacts with another individual from group \(j\) is enough to
set the interaction \(A_{i,j}\) to 1. This simplified representation of
food webs is a highly valuable source of ecological information (Pascual
\emph{et al.} 2006) even though it overlooks important factors regarding
interaction strengths. These, in turn, can be represented using weighted
interactions, which better describe the energy flows, demographic
impacts or frequencies of interactions between nodes (Berlow \emph{et
al.} 2004; Borrett \& Scharler 2019), with \(A_{i,j} \in \mathbb{N}\) or
\(\mathbb{R}\) depending on the measure. For example, they can be used
to estimate the average number of prey individuals consumed by the
predators in a given time period. This extra amount of ecological
information typically comes at a cost of greater sampling effort or data
requirement in predictive models (Strydom \emph{et al.} 2021), which can
lead to high uncertainties when building these types of networks.

The uncertainty and spatiotemporal variability of both types of biotic
interactions (Boolean and weighted) can be represented
probabilistically. On one hand, Boolean interactions follow a Bernoulli
distribution \(A_{i,j} \sim {\rm Bernoulli}(p)\), with \(p\) being the
probability of interactions. The only two possible outcomes are the
presence (\(A_{i,j} = 1\)) or absence (\(A_{i,j} = 0\)) of an
interaction between the two nodes. Weighted interactions, on the other
hand, can follow various probability distributions depending on the
measure used. In this case, the event's outcome is the value of
interaction strength. For instance, weights can follow a Poisson
distribution \(A_{i,j} \sim {\rm Poisson}(\lambda)\) when predicting
frequencies of interactions between pairs of nodes, with \(\lambda\)
being the expected rate of interaction. The definition and
interpretation of parameters like \(p\) and \(\lambda\) are inextricably
linked to environmental and biological factors such as species relative
abundance, traits, area, and time, depending on the type of interaction.
Because Boolean species interactions are much more documented in the
literature, our primary focus in this contribution will be on addressing
the challenges in defining and interpretating \(p\) for pairwise species
interactions.

The first aspect to take into consideration when estimating or using
probabilities of interactions is knowing if they describe the likelihood
of potential or realized interactions. A potential interaction is
defined as the biological capacity of two species to interact (i.e., the
probability that they \emph{can} interact) whereas a realized
interaction refers to the materialization or observation of this
interaction in a delineated space and time period (i.e., the probability
that they \emph{will} interact). Here, we will use the terms
\emph{metaweb} to designate networks of potential interactions and
\emph{local networks} for those of realized interactions. Frequent
confusion arises among ecologists over the use of these two terms,
especially in a probabilistic context. Indeed, in many studies of
probabilistic ecological networks, it remains unclear when authors
describe potential or realized interactions, or when so-called
probabilities are actually \emph{interaction scores}. A better
understanding of these differences would alleviate interpretation errors
and help ecologists use these numbers more appropriately.

@@

\begin{itemize}
\tightlist
\item
  It means the probability that two taxa can interact, regardless of
  biological plasticity, environmental variability, or co-occurrence.
\item
  One observation is enough to set this probability to one.
\item
  Can we turn this into a local network realisation that is also
  probabilistic and intuitive?
\item
  Low probability with rare interactions?
\end{itemize}

@@

\leavevmode\hypertarget{ref-Berlow2004IntStr}{}%
Berlow, E.L., Neutel, A.-M., Cohen, J.E., De Ruiter, P.C., Ebenman, B.,
Emmerson, M., \emph{et al.} (2004). Interaction strengths in food webs:
Issues and opportunities. \emph{Journal of Animal Ecology}, 73,
585--598.

\leavevmode\hypertarget{ref-Borrett2019WalPar}{}%
Borrett, S.R. \& Scharler, U.M. (2019). Walk partitions of flow in
Ecological Network Analysis: Review and synthesis of methods and
indicators. \emph{Ecological Indicators}, 106, 105451.

@@

\leavevmode\hypertarget{ref-Elton2001AniEco}{}%
Elton, C.S. (2001). \emph{Animal Ecology}. University of Chicago Press,
Chicago, IL.

@@

\leavevmode\hypertarget{ref-Strydom2021Roadmapa}{}%
Strydom, T., Catchen, M.D., Banville, F., Caron, D., Dansereau, G.,
Desjardins-Proulx, P., \emph{et al.} (2021). A roadmap towards
predicting species interaction networks (across space and time).
\emph{Philosophical Transactions of the Royal Society B-Biological
Sciences}, 376, 20210063.
